---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Loading data:
```{r loaddata, echo=TRUE}
data <- read.csv("activity.csv",header = TRUE)
```

head(data):
```{r showdata, echo=TRUE}
head(data)
```


## What is mean total number of steps taken per day?
```{r stepsByDay, echo=TRUE}
stepsByDay <- aggregate(steps~date, data, sum)
head(stepsByDay)
```

Histogram:
```{r histogram1, echo=TRUE}
hist(stepsByDay$steps, col = "chocolate"
     , xlab = "Steps", main = "Total Number of Steps Per Day")
```

Mean & Median:
```{r mean1, echo=TRUE}
mean1 <- mean(stepsByDay$steps)
mean1
median1 <- median(stepsByDay$steps)
median1
```

## What is the average daily activity pattern?

```{r averageSteps, echo=TRUE}
averageSteps <- aggregate(steps~interval, data, mean)
head(averageSteps)
```

Time Series Plot:
```{r plot1, echo=TRUE}
with(averageSteps, plot(interval, steps,
                        type="l",
                        main="Average number of steps taken in 5-min interval"
                , col = "blue")) 
```

Max data is:
```{r max, echo=TRUE}
max <- max(averageSteps$steps)
max
```

The interval contains max data is:
```{r intervalmax, echo=TRUE}
subsetData <- subset(averageSteps, steps == max)
subsetData$interval
```

## Imputing missing values
Total number of rows with NAs:
```{r missing, echo=TRUE}
colSums(is.na(data))
```

Average steps in each interval:
```{r newsteps, echo=TRUE}
data$newSteps <- averageSteps$steps
head(data)
```

Filling Nas/Imputing:
```{r fill, echo=TRUE}
data$steps[is.na(data$steps)] <- data$newSteps
head(data)
```

New data set:
```{r newdata, echo=TRUE}
data$newSteps <- NULL
newdata <- data
head(newdata)
```

```{r sumdata, echo=TRUE}
sumdata <- aggregate(steps~date, newdata, sum)
head(sumdata)
```

Histogram New:
```{r histogram2, echo=TRUE}
hist(sumdata$steps, col = "blue"
     , xlab = "Steps"
     , main = "Total Number of Steps Per Day (with the missing data filled in)")
```

Mean & Median:
```{r mean2, echo=TRUE}
mean2 <- mean(sumdata$steps)
mean2
median2 <- median(sumdata$steps)
median2
```

Difference between original and imputed:
```{r diff, echo=TRUE}
mean2 - mean1
median2 - median1
```

## Are there differences in activity patterns between weekdays and weekends?
Time Package:
install.packages("timeDate")
```{r load, echo=TRUE}
library(timeDate)
```

Checking for weak day:
```{r check, echo=TRUE}
newdata$Weekday <- isWeekday(newdata$date)
head(newdata)
```

Average steps:
Weekday:
```{r weekday, echo=TRUE}
weekday <- subset(newdata, newdata$Weekday == "TRUE")
weekdayMean <- aggregate(steps ~ interval, data = weekday, mean)
head(weekdayMean)
```

Weekend:
```{r weekend, echo=TRUE}
weekend <- subset(newdata, newdata$Weekday == "FALSE")
weekendMean <- aggregate(steps ~ interval, data = weekend, mean)
head(weekendMean)
```

Panel Plot:
```{r panel, echo=TRUE}
layout(matrix(c(1,1,2,2), 2, 2, byrow = TRUE))
plot(weekdayMean$interval, weekdayMean$steps
     , xlab = "interval", ylab = "Number of steps"
     , main ="Weekday", col ="blue", type="l") 
plot(weekendMean$interval, weekendMean$steps
     , xlab = "interval", ylab = "Number of steps"
     , main ="Weekend", col ="red", type="l")
```
